---
title: "BA888-Feb 13"
author: "Jiao Sun"
date: "2/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(jsonlite)
library(tidytext)
library(stringr)
library(ggplot2)
library(dplyr)
library(ggthemes)
```

### Business Porblem   
-Finding restaurants through directory services apps and review forum apps have become a crucial part of business discovery. It has changed the way restaurants of different tiers attain brand loyalty and market themselves.   
-Yelp’s terabytes of historical data on restaurants’ reviews, check-ins, and special attributes have led to the ability to uncover valuable trends to aid businesses in understanding their impact on customers and what make customers tick.   
-How can a restaurant leverage information on Yelp to increase their chances of attaining their business goals and preventing permanent closure? What aspects make customers choose a restaurant over its competitors? 

#################### sample ####################
```{r}
yelp = stream_in(file("business.json"))
yelp_flat = flatten(yelp)
yelp_buss = as.data.frame(yelp_flat)

write_csv(yelp_buss, "yelpbusiness.csv")
data = read_csv("yelpbusiness.csv")

data1 = data[1:500, ]
df1 = data1 %>%
 select(1:12) %>%
 mutate(categories = str_to_lower(categories)) %>%
 filter(str_detect(categories, "restaurants"))

df2 = data1 %>%
 select(1:12) %>%
 mutate(categories = str_to_lower(categories))

df3 = anti_join(df2, df1)
df1_1 = df3 %>%
  filter(str_detect(categories,"food"))
df1_2 = rbind(df1, df1_1)

## check the restaurant businesses
df3_1 = anti_join(df2, df1_2)
df4 = df3_1 %>%
  separate(categories, into = paste0("v", 1:100))
df4 = df4[colSums(!is.na(df4)) > 0] 
words = df4 %>%
  pivot_longer(cols = v1:v24,
               names_to = "pos",
               values_to = "token",
               values_drop_na = TRUE) %>% 
  filter(str_length(token) > 0) %>%
  group_by(token) %>%
  count(sort = T)

## combine restaurants to the original dataset
data1_1 = data1 %>%
  mutate(categories = str_to_lower(categories))
newdata1 = inner_join(data1_1, df1_2)
```

```{r}
## filter business which has "restaurants" and "food" in the categories ##
df = data %>%
  select(1:12) %>%
  filter(str_detect(categories, "restaurants"))

data_check = data %>%
  select(1:12)
df_check = anti_join(data_check, df)

df_food = df_check %>%
  filter(str_detect(categories, "food"))
df_new = rbind(df, df_food)  

df_check1 = anti_join(data_check, df_new)

words_left = df_check1 %>%
  select(categories) %>%
  separate(categories, into = paste0("v", 1:100)) %>%
  pivot_longer(cols = v1:v100,
               names_to = "pos",
               values_to = "token",
               values_drop_na = TRUE) %>% 
  filter(str_length(token) > 0) %>%
  group_by(token) %>%
  count(sort = T)

## Only RESTAURANT Now ##
restaurants = inner_join(data, df_new)

## delet the coloumn that have more than 50% NAs.
restaurants = restaurants[colSums(!is.na(restaurants)) > 0.5*nrow(restaurants)]

## only open restaurants ##
open = restaurants %>%
  filter(is_open == 1)

## write out the clean dataset
write_csv(restaurants, "res.csv")
```

#################### working on restaurant ##################
```{r warning=FALSE}
res <- read.csv("restaurants.csv")
res_in_state = res %>%
  mutate(x = 1) %>%
  select(x, name, state, stars, review_count, is_open, 
         attributes.RestaurantsPriceRange2) %>%
  na.omit(.) %>%
  group_by(state) %>%
  summarise(num = sum(x)) %>%
  filter(num > 50)

is.data.frame(res_in_state)

stars_by_state = res %>%
  select(name, state, stars, review_count, is_open, 
         attributes.RestaurantsPriceRange2) %>%
  na.omit(.) %>%
  group_by(state) %>%
  summarise(stars = mean(stars))

## Visualization
ggplot(res, aes(x=is_open)) +
  geom_histogram(stat="count",fill="#FF6677") +
  labs(title="Open vs Closed Restuarants",x="Open?")+
  xlim(c("No","Yes")) +
  ylim(0, 58000) +
  theme_economist() +
  scale_color_economist() +
  theme(axis.text.x = element_text(hjust = 15.5))

ggplot(res, aes(x=stars)) +
  geom_histogram(stat="count",position="dodge2",aes(fill=factor(is_open))) +
  labs(title="The Distribution of Star Ratings",x="star ratings",fill="Open?") +
  scale_fill_discrete(labels=c("No","Yes")) +
  theme_economist() + 
  scale_color_economist()
```


