---
title: "Project Overview - Can Yelp Reviews Save the Restaurant Industry?"
author: "Xinman Liu, Melissa Putur, Jiao Sun, Adil Bin Wahab, Rui Xu"
date: "2/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(jsonlite)
library(tidytext)
library(stringr)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(readr)
library(tidyr)
```

```{r message=FALSE, warning=FALSE}
#Loading the datasets
#restaurants <- read_csv("/Users/Adil/Documents/RStudio/cananda_restaurant.csv") #adil file path
restaurants <- read_csv("../Classes/888/Canada CSVs/cananda_restaurant.csv") #missy file path
#dim(restaurants)
#reviews <- read_csv("/Users/Adil/Documents/RStudio/capstoneData/canada_reviews.csv") #adil file path
reviews <- read_csv("../Classes/888/Canada CSVs/canada_reviews.csv") #missy file path
#checkIns <- read_csv("/Users/Adil/Documents/RStudio/capstoneData/canada_checkin.csv")
checkin <- read_csv("../Classes/888/Canada CSVs/canada_checkin.csv")
```


### Business Problem   
Finding restaurants through directory services apps and review forum apps have become a crucial part of business discovery. It has changed the way restaurants of different tiers attain brand loyalty and market themselves.   

Yelp’s terabytes of historical data on restaurants’ reviews, check-ins, and special attributes have led to the ability to uncover valuable trends to aid businesses in understanding their impact on customers and what make customers tick.   

How can a restaurant leverage information on Yelp to increase their chances of attaining their business goals and preventing permanent closure? What aspects make customers choose a restaurant over its competitors? 


### Ways to Solve Our Problem
**1. descriptive + diagnostic analytics**
Study the most essential aspects of a restaurant based on cuisine and price level to understand what most customers value (e.g. Do customers at low-price restaurants value the same “things” as mid-price and high-price restaurants).
**2. predictive analytics**
Predict whether a restaurant will close in the near future based on specific variables.
**3. prescriptive Analytics**
Create a recommendation system that predicts users’ preferences based on their reviews. prescriptive Analytics


### Data Source   
We got our data from [Kaggle](https://www.kaggle.com/yelp-dataset/yelp-dataset), and there are 4 datasets in total, namely, business, check-in, reviews, and users. 
**Business dataset** includes business types, attributes, location, review count, star ratings, and open hours.
**User dataset** about reviews on each business: text, rating, review count, and reactions (useful, cool, funny).
**Check-ins datasets** includes the exact timestamp a user checked into a business.
**Review dataset** includes the review text of each business with user ids, usefulness count, and reactions (useful, cool, funny).


### Data Cleaning
**Business Dataset:** For the project we are only interested in exploring reviews, star ratings, and interactions with restaurants. The business dataset contains information for all business reviewed by yelp. To narrow down this dataset we manually reviewed the top 1,000 phrases that businesses were tagged with to identify phrases that would be a good fit for restaurant. We filtered the complete list to only include businesses that were tagged with the phrases “food” and/or “restaurant”. Using these phrases we were able to identify ~75,000 restaurants. 
**Review Dataset:** The review dataset contains reviews for all businesses on Yelp. We reduced the review dataset to only include reviews for restaurants (as identified by the above processing steps). After this reduction the dataset contained over 4 million reviews.
**Check-In Dataset:** The check-in dataset was structured so that each business was represented as a row with a comma separated list of check-ins by date. To make the data easier to work with we rearranged it so that each check-in was represented as a row rather than a list.
**Final Data Cleaning Criteria:** To balance having a robust datset that minimizes long compute time while preserving data quality, the dataset was filtered to only include Canadian cities (Toronto, Mississauga and Calgary). Montréal was excluded as a majority of the reviews are in French. Restuarants with less than 10 reviews were dropped, and only the reviews of users who have written at least 10 reviews were included. This criteria was implemented to reduce the frequency of biased reviews and create a fair comparison between the restaurants.

**Post-cleaned Dataset Information**: 
Table 1: Number of reviews of each resturant (`nrow(reviews)` rows) - 200 KB (CSV)
Table 2: Restaurant general info and attributes (`nrow(restaurants)` rows) - reduced to 4 MB (CSV) from 34 MB (JSON)
Table 3: Restuarant check-ins (`nrow(checkin)` rows) - reduced to 38.5 MB (CSV) from 390 MB (JSON)
Table 4: Restaurant reviews (371,610 rows) - reduced to 320 MB (CSV) from 4.98 GB (JSON)

###Exploratory Data Analysis

##Restaurants Exploratory Data Analysis 

There are 1,422 restaurants in Calgary, 808 restaurants in Mississauga, Toronto has the most restaurants with 5,547.
```{r}
# restaurants are in each city
#restaurants %>% group_by(city) %>% count(city)
restaurants %>% 
  ggplot(aes(x=city))+
  geom_bar(fill=c("#F15C4F", "#FCD6D3","#F8ADA8"))+
  theme_classic()+
  labs(title = "Restaurant Distribution by City",
       y="Amount")
```

Approximately 6,000 restaurants in the dataset are open and approximately 2,000 restaurants have closed. 
```{r}
open <- restaurants %>% group_by(is_open) %>% count(is_open)
open[open$is_open==0,1]="no"
open[open$is_open==1,1]="yes"
open
open %>% 
  ggplot(aes(is_open,n))+
  geom_col(fill=c("#F15C4F", "#F8ADA8"))+
  labs(title = "Current Restuarant Status",
       x="is open?",
       y="Amount")+
  theme_classic()
```

The majority of the restaurants in the dataset are in the "2-Dollar-Sign" price range. 
```{r message=FALSE, warning=FALSE}
restaurants %>% 
  ggplot(aes(attributes_RestaurantsPriceRange2))+
  geom_histogram(binwidth=1,fill="#F15C4F",color="#4D4D4D")+
  labs(title = "Distribution of Price Range",
       x="Price Range")+
  theme_classic()
```

##Reviews Exploratory Data Analysis
The dataset contains ~ 372,000 reviews for the 7,777 restaurants. On average, each restaurant has 48 reviews. 
```{r message=FALSE, warning=FALSE}
print(paste0("number of reviews: ", nrow(reviews)))
print(paste0("average # of reviews: ", round(nrow(reviews)/nrow(restaurants))))
#summary(reviews)
```

#shouldn't be run? same as other graph but with different bins?
```{r}
review_length <- sapply(strsplit(reviews$text, " "), length)
review_length <- as.data.frame(review_length)

# Dist count
ggplot(review_length) + geom_histogram(aes(x=review_length),bins=25,
                                       fill="#D32323") +
  labs(title="Distribution of Review Length",x="review length") +
  theme_minimal()
```

The distribution of the review length is right-skewed with a mode of around 30 to 60 words, which represents the word count of over 20% of all reviews. More than 50% of all the reviews are between 30 to 120 words. The average review length is around 130 words while the median is around 100 words. 
```{r}
# Dist freq 
ggplot(data=review_length, aes(x=review_length)) +
geom_histogram(aes(y = stat(count) / sum(count)),
               bins = 30, fill="#D32323") +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Distribution of Review Length",
       x="Review Length",y="Relative Frequency") +
  theme_minimal() + scale_x_continuous(breaks=seq(0,1000,60))
```

```{r}
print(paste0("mean ", (round(mean(review_length$review_length)))))
print(paste0("median ", (median(review_length$review_length))))
```

## Check-Ins Exploratory Data Analysis:


```{r}
checkin$business_id=as.character(checkin$business_id)
restauranu_count=checkin %>% group_by(business_id) %>% count(.)

ggplot(checkin, aes(x=business_id)) +
  geom_histogram(stat="count",fill="#D32323") +
  labs(title="Distribution of check-in count")+
  theme_classic()
```

```{r}
### Checkins by Star Ratings
restaurants$business_id=as.character(restaurants$business_id)
new_table=left_join(x=checkin,y=restaurants)
checkin_count=new_table %>% group_by(stars) %>% count(.)
glimpse(new_table)
ggplot(new_table, aes(x=stars)) +
  geom_histogram(stat="count",fill="#D32323") +
  labs(title="Checkins by Star Ratings")+
  theme_classic()
##4 stars have the largest numbers of checkins, followed by 3.5 and 3. While 5 and 1 stars
##have the smallest number of checkins maybe because it is too expensive.
```

```{r}
### Distribution of star ratings
ggplot(restaurants)+
  geom_bar(aes(x=city,fill=factor(stars)),position="dodge")+
  labs(title="Distribution of star ratings in Different City")
##Toronto has the largest number of hotels, and Toronto also has the largest number of 5, 4 
##and 3 stars resutaurants.
```

##Check-In EDA
```{r}
checkinsforchart <- checkins %>% 
  select(month_year) %>% 
  group_by(month_year) %>% 
  add_count() %>%
  distinct()

checkinsforchart %>% 
  ggplot() +
  geom_line(aes(x = month_year, y = n, group = 1), color = "#D32323", size = 2) + 
  theme_minimal()+
  theme(axis.text.x=element_text(angle=90, hjust=1)) +
  labs(title = "Check-In by Day", x = "date", y="count")
```




```{r}
#res <- read.csv("restaurants.csv") #original file path
res <- read.csv("../Classes/888/Canada CSVs/cananda_restaurant.csv") #missy file path 
res_in_state = res %>%
  mutate(x = 1) %>%
  select(x, name, state, stars, review_count, is_open, 
         attributes.RestaurantsPriceRange2) %>%
  na.omit(.) %>%
  group_by(state) %>%
  summarise(num = sum(x)) %>%
  filter(num > 50)

is.data.frame(res_in_state)

stars_by_state = res %>%
  select(name, state, stars, review_count, is_open, 
         attributes.RestaurantsPriceRange2) %>%
  na.omit(.) %>%
  group_by(state) %>%
  summarise(stars = mean(stars))

## Visualization
ggplot(res, aes(x=is_open)) +
  geom_histogram(stat="count",fill="#FF6677") +
  labs(title="Open vs Closed Restuarants",x="Open?")+
  xlim(c("No","Yes")) +
  ylim(0, 58000) +
  theme_economist() +
  scale_color_economist() +
  theme(axis.text.x = element_text(hjust = 15.5))

ggplot(res, aes(x=stars)) +
  geom_histogram(stat="count",position="dodge2",aes(fill=factor(is_open))) +
  labs(title="The Distribution of Star Ratings",x="star ratings",fill="Open?") +
  scale_fill_discrete(labels=c("No","Yes")) +
  theme_economist() + 
  scale_color_economist()

restaurant1=res %>% select(state,stars)
restaurant1 =restaurant1[complete.cases(restaurant1),]
restaurant2 = res[complete.cases(res),]
names(res)
ggplot(res, aes(attributes.GoodForKids))+
  geom_histogram(stat="count",fill="red")+
  facet_grid(is_open~.)
ggplot(res,aes(attributes.RestaurantsPriceRange2,fill=factor(attributes.RestaurantsReservations),stat="identity"))+
  geom_bar()+
  labs(title = "Restaurants Reservations by price range",
       x="Resaurant price range", 
       fill="Resaurant Reservations")
ggplot(restaurant1)+
  geom_bar(aes(x=state,fill=factor(stars)),position="dodge")+
  labs(title="Distribution of Star Restaurants in the State",
       y="The number of each stars")
ggplot(restaurant1,aes(x=state,y=stars))+
  stat_summary(fun.y = "mean",geom="bar")+
  labs(title="Average star rating by state",
       y="Avergae star rating")
ggplot(res)+
  geom_bar(aes(x=attributes.RestaurantsPriceRange2,fill=factor(stars)),position="dodge")+
  labs(title="Price Range for Star Restaurants",
       x="Resaurant Price Range",
       y="The number of each stars")

```

```{r}
yelp = stream_in(file("business.json"))
yelp_flat = flatten(yelp)
yelp_buss = as.data.frame(yelp_flat)

write_csv(yelp_buss, "yelpbusiness.csv")
data = read_csv("yelpbusiness.csv")

data1 = data[1:500, ]
df1 = data1 %>%
 select(1:12) %>%
 mutate(categories = str_to_lower(categories)) %>%
 filter(str_detect(categories, "restaurants"))

df2 = data1 %>%
 select(1:12) %>%
 mutate(categories = str_to_lower(categories))

df3 = anti_join(df2, df1)
df1_1 = df3 %>%
  filter(str_detect(categories,"food"))
df1_2 = rbind(df1, df1_1)

## check the restaurant businesses
df3_1 = anti_join(df2, df1_2)
df4 = df3_1 %>%
  separate(categories, into = paste0("v", 1:100))
df4 = df4[colSums(!is.na(df4)) > 0] 
words = df4 %>%
  pivot_longer(cols = v1:v24,
               names_to = "pos",
               values_to = "token",
               values_drop_na = TRUE) %>% 
  filter(str_length(token) > 0) %>%
  group_by(token) %>%
  count(sort = T)

## combine restaurants to the original dataset
data1_1 = data1 %>%
  mutate(categories = str_to_lower(categories))
newdata1 = inner_join(data1_1, df1_2)

```


```{r}
## filter business which has "restaurants" and "food" in the categories ##
df = data %>%
  select(1:12) %>%
  filter(str_detect(categories, "restaurants"))

data_check = data %>%
  select(1:12)
df_check = anti_join(data_check, df)

df_food = df_check %>%
  filter(str_detect(categories, "food"))
df_new = rbind(df, df_food)  

df_check1 = anti_join(data_check, df_new)

words_left = df_check1 %>%
  select(categories) %>%
  separate(categories, into = paste0("v", 1:100)) %>%
  pivot_longer(cols = v1:v100,
               names_to = "pos",
               values_to = "token",
               values_drop_na = TRUE) %>% 
  filter(str_length(token) > 0) %>%
  group_by(token) %>%
  count(sort = T)

## Only RESTAURANT Now ##
restaurants = inner_join(data, df_new)

## delet the coloumn that have more than 50% NAs.
restaurants = restaurants[colSums(!is.na(restaurants)) > 0.5*nrow(restaurants)]

## only open restaurants ##
open = restaurants %>%
  filter(is_open == 1)

## write out the clean dataset
write_csv(restaurants, "res.csv")
```






