---
title: "CohortAnalysis.rmd"
author: "Rui Xu"
date: "4/15/2020"
output: html_document
---
#Cohort Analysis
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(reshape2)
library(zoo)
options(scipen=999)

reviews = read_csv('~/Desktop/Capstone/canada_reviews_ms.csv')
```
## Consumer retention
```{r}
data = reviews %>%
  mutate(year = year(yelping_since),
         user_id = factor(user_id)) %>%
  filter(year >= 2015) %>%
  select(user_id, date, yelping_since, review_id)
```
```{r}
data$review_yr_q = format(as.yearqtr(data$date,"%Y-%m-%d"), format="%y/0%q") #customer review period.
data$CohortGroup = format(as.yearqtr(data$yelping_since, '%Y-%m-%d'), format="%y/0%q") # #user create account period
```
```{r}
CohortAnalysis = data %>% 
  group_by(CohortGroup,review_yr_q) %>%            
  summarize(user_id  = n_distinct(user_id),
            review_id = n_distinct(review_id)) %>% #calculate unique user num by each review period and yelp start period.
  rename(TotalUsers= user_id,
         Totalreviews = review_id) %>%
  mutate(ReviewPerPerson = Totalreviews/TotalUsers) %>%
  arrange(CohortGroup,review_yr_q) %>% #assign number to each review period
  ungroup() %>%
  group_by(CohortGroup) %>%               
  mutate(CohortPeriod =row_number())
```
```{r}
CA_user_group_size = CohortAnalysis %>%
  filter(CohortPeriod == 1) %>%
  select(CohortGroup,review_yr_q,TotalUsers) 

user_retention = CohortAnalysis %>% 
  select(CohortGroup,CohortPeriod,TotalUsers) %>%
  spread(CohortGroup,TotalUsers) # convert long table to wide table.

user_retention[,-1] <- user_retention[,-1] %>% 
  t() %>%
  `/`(CA_user_group_size$TotalUsers) %>% t() %>% as.data.frame() # Calculate the percentage of user num to each group's total num.
```
```{r}
# Visualization
user_retentionT <- t(user_retention) %>% 
  .[2:nrow(.),]  %>% 
  as.data.frame()

user_retentionT$CohortPeriod <- row.names(user_retentionT)
row.names(user_retentionT) <- NULL
user_retentionT <- user_retentionT[,c(17,1:16)]
user_retentionT1 <- user_retentionT %>% 
  melt(id.vars = 'CohortPeriod',
       variable.name = 'CohortGroup',
       value.name = 'TotalUsers')
```

```{r}
library("Cairo")
library('sysfonts')
library('showtextdb')
library("showtext")

ggplot(user_retentionT1 ,aes(CohortGroup,CohortPeriod,fill=TotalUsers))+
  geom_tile(colour='white') +
  geom_text(aes(label = ifelse(TotalUsers !=0,paste0(round(100*TotalUsers,2),'%'),'')),
            colour = 'black', size = 2.3) +
  scale_fill_gradient2(limits=c(0,.8),
                       low="#F4D03F", 
                       mid ='#FADBD8', 
                       high="#CB4335",
                       midpoint = median(user_retentionT1$TotalUsers, na.rm =TRUE),na.value = "grey90") +
  scale_y_discrete(limits = rev(unique(user_retentionT1$CohortPeriod))) +
  scale_x_discrete(position = "top") +
  labs(title='Cohort Analysis - User Quarterly Retention',
       subtitle ='2018 new customers performed less retention than the past customers.') +
  theme(
    legend.position = "bottom")
```
## Review count
```{r}
CA_review_group_size = CohortAnalysis %>%
  filter(CohortPeriod == 1) %>%
  select(CohortGroup,review_yr_q,Totalreviews) 

review_posted = CohortAnalysis %>% 
  select(CohortGroup,CohortPeriod,Totalreviews) %>%
  spread(CohortGroup,Totalreviews) # convert long table to wide table.

review_postedT = t(review_posted) %>% 
  .[2:nrow(.),]  %>% 
  as.data.frame()

review_postedT$CohortPeriod <- row.names(review_postedT) # reshape wide table, then turn back to long to do heat map.
row.names(review_postedT) <- NULL
review_postedT <- review_postedT[,c(17,1:16)] 
review_postedT1 <- review_postedT %>% 
  melt(id.vars = 'CohortPeriod',
       variable.name = 'CohortGroup',
       value.name = 'TotalReviews')
```
```{r}
# Visualization
t=max(review_postedT1$TotalReviews)
ggplot(review_postedT1 ,aes(CohortGroup,CohortPeriod,fill=TotalReviews))+
  geom_tile(colour='white') +
  geom_text(aes(label = ifelse(TotalReviews !=0,TotalReviews,"")),
            colour = 'black', size = 2.7) +
  scale_fill_gradient2(limits=c(0,t),
                       low="#F4D03F", 
                       mid ='#FADBD8', 
                       high="#CB4335",
                       midpoint = median(review_postedT1$TotalReviews, na.rm =TRUE),na.value = "grey90") +
  scale_y_discrete(limits = rev(unique(review_postedT1$CohortPeriod))) +
  scale_x_discrete(position = "top") +
  labs(title='Cohort Analysis - Quarterly Reviews',
       subtitle = 'Reviews by new users trend to decrease from 2015 to 2018.')
```
```{r}
# Cohort Curve
review_postedT2 = review_postedT1 %>%
  filter(CohortGroup == 'V1'|CohortGroup == 'V2'|CohortGroup == 'V3')
ggplot(review_postedT2,aes(CohortPeriod,TotalReviews)) +
     geom_line(aes(group = CohortGroup, colour = CohortGroup))+
     scale_color_manual(name="Time group",
                       labels=c("1st quarter","2nd quarter","3rd quarter"),
                       values = c("#CB4335", "#EC7063", '#F5B7B1')) +
     labs(title = 'New User Reviews in First Three Quarters',
          subtitle = 'First quarter reviews trend to decrease from 2015 to 2018.',
          x='Cohort period',
          y='Total reviews') +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent",colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA)
  )
  
```


# Review pre person
```{r}
CA_reviewprp_size = CohortAnalysis %>%
  filter(CohortPeriod == 1) %>%
  select(CohortGroup,review_yr_q,ReviewPerPerson) 

review_per_person = CohortAnalysis %>% 
  select(CohortGroup,CohortPeriod,ReviewPerPerson) %>%
  spread(CohortGroup,ReviewPerPerson) # convert long table to wide table.

review_per_personT = t(review_per_person) %>% 
  .[2:nrow(.),]  %>% 
  as.data.frame()

review_per_personT$CohortPeriod <- row.names(review_per_personT) # reshape wide table, then turn back to long to do heat map.
row.names(review_per_personT) <- NULL
review_per_personT = review_per_personT[,c(17,1:16)] 
review_per_personT1 <- review_per_personT %>% 
  melt(id.vars = 'CohortPeriod',
       variable.name = 'CohortGroup',
       value.name = 'ReviewPerPerson')
```
```{r}
#Visualization
t1=max(review_per_personT1$ReviewPerPerson)
ggplot(review_per_personT1,aes(CohortGroup,CohortPeriod,fill=ReviewPerPerson))+
  geom_tile(colour='white') +
  geom_text(aes(label = ifelse(ReviewPerPerson !=0,round(ReviewPerPerson,2),"")),
            colour = 'black', size = 3) +
  scale_fill_gradient2(limits=c(1,t1),
                       low="#F8F9F9", 
                       mid ='#CCD1D1', 
                       high="#34495E",
                       midpoint = median(review_per_personT1$ReviewPerPerson, na.rm =TRUE),na.value = "grey90") +
  scale_y_discrete(limits = rev(unique(review_per_personT1$CohortPeriod))) +
  scale_x_discrete(position = "top") +
  labs(title='Cohort Analysis - Quarterly Reviews Per Person',
       subtitle = "Users who stay until 2nd quarter frequently reviewing during this period.")
```
```{r}
# Cohort Curve
review_per_personT2 = review_per_personT1 %>%
  filter(CohortGroup == 'V1'|CohortGroup == 'V2'|CohortGroup == 'V3')
ggplot(review_per_personT2,aes(CohortPeriod,ReviewPerPerson)) +
     geom_line(aes(group = CohortGroup, colour = CohortGroup))+
     scale_color_manual(name="Time group",
                       labels=c("1st quarter","2nd quarter","3rd quarter"),
                       values = c("#707B7C", '#EB984E', "#95A5A6")) +
     labs(title = 'Reviews Per Returned User in First Three Quarter',
          subtitle = 'Users who stay until 2nd quarter frequently reviewing during this period',
          x='Cohort period',
          y='Review per person') +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent",colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA)
  )
```

**Takeaway**
*User Rentention* 
  - During 2015 to 2017, over 40% new user would return to review restaurant on Yelp within two quarters, over 30% users keep reviewing on yelp after one year. After two years, the retention rate went to 20% or less. 
  - However, user retention in 2018 4th quarter drop a lot. Around 15% users come back to Yelp review in this 4th quater only, even the new users who start to use Yelp within one year. (Could it due to data loss?)   
*User Review posted* New users came in 2015 and 2016 are more likely contributed to Yelp content in first quarter than new users came in after 2016.
*Average review posting* The second quarter is the core user retention period. Users not only keep relativly high rentetion rate, but also likely to post more reviews in this period.
