---
title: "Recommender"
author: "Rui Xu"
date: "4/16/2020"
output: pdf_document
---
```{r}
library(tidyverse)
library(lubridate)
library(reshape2)
library(zoo)
options(scipen=999)

library(recommenderlab) # for recommender

reviews = read_csv('~/Desktop/Capstone/canada_reviews_ms.csv')
restaurant = read.csv('~/Desktop/Capstone/canada_restaurants_ms_edits.csv')
```
```{r}
# Pick restaurants into recommender sample
restaurant %>%
  filter(is_open == 1) %>%
  group_by(city) %>%
  summarise(num_rest = sum(n()),
            num_review = sum(review_count))
```
I pick open restaurants in Mississauga where small number of business exists, in case further technique problem when manipulate data structure.
```{r message=FALSE, warning=FALSE}
RecomRest = restaurant %>%
  filter(is_open == 1,
         city == 'Mississauga') %>%
  select(business_id, rest_name = name) %>%
  inner_join(.,reviews, by='business_id')
```
```{r}
# Inspect whether reviewers who post less would have a different rating scheme
ggplot(RecomRest, aes(y=average_stars,x=review_count)) +
  geom_point(size=0.8) +
  xlim(c(0,50))
```
```{r}
ggplot(RecomRest, aes(y=average_stars,x=review_count)) +
  geom_point(size=0.8) +
  xlim(c(600,5000))
```

Pre-process Data
```{r}
# reassign unique user_id and unique business_id
userId = RecomRest %>%
  select(user_id) %>%
  unique() %>%
  mutate(userIdR = row_number())
bussId = RecomRest %>%
  select(business_id) %>%
  unique() %>%
  mutate(bussIdR = row_number())
```
```{r}
# keep varibles that used for recommendation system: users id, star rating, business id
tmp = inner_join(userId, RecomRest, 'user_id')
RecomData = inner_join(tmp, bussId, 'business_id')
RecomData1 = RecomData %>%
  select(userIdR,stars,bussIdR,date)
RecomData1
RecomData1$userIdR = as.character(RecomData$userIdR)
RecomData1$userIdR = as.numeric(RecomData1$userIdR)
RecomData1$bussIdR = as.character(RecomData1$bussIdR)
RecomData1$bussIdR = as.numeric(RecomData1$bussIdR)
RecomData1$date = as.numeric(as.POSIXct(RecomData1$date))
```


```{r}
RecomData1
```
```{r}
write_csv(RecomData1, 'RecomData.csv')
df = read_csv('RecomData.csv')
```
```{r}
glimpse(df)
```
```{r}
# Exist one userId rating one bussId multiple times, we select the least date one as final stars.
df1 = df%>%
  group_by(userIdR,bussIdR) %>%
  filter(row_number() == 1) %>%
  select(-date)
```
```{r}
## put the dataset into a wide format
ratingR = df1 %>%
  pivot_wider(names_from = bussIdR,
              values_from = stars,
              names_prefix = 'r')

ratingR[1:5, 1:5]
```
```{r}
dim(ratingR)
```
```{r}
## convert the dataframe into a matrix to retain user ids
rownames(ratingR) =  ratingR$userIdR
ratingR$userIdR = NULL
rat_mat = as.matrix(ratingR)
rownames(rat_mat)[1:5]
colnames(rat_mat)[1:5]
```
```{r}
## convert to realRatingMatrix from package
## use that matrix to create a testing scheme
rat_mat = as(rat_mat, "realRatingMatrix")
SAMPLE = evaluationScheme(rat_mat, 
                          method="split", 
                          train=.5, 
                          given=1)
class(SAMPLE)
SAMPLE
```
```{r}
# Model fitting
reco1 = Recommender(getData(SAMPLE, "train"), 
                    method="UBCF", # User-Based Collaborative filtering - utilizes user similarities
                    param = list(normalize = NULL, # do not normalize the ratings
                                 method = "Cosine")) # use cosine similiarty on the ratings

class(reco1) 
reco1 # learned using 6903 users.
getModel(reco1)
```

```{r}
# Look at the prediction
N_PREDS = 10 # set the number of recommendations in the top-N list
pred1 = predict(reco1, getData(SAMPLE, "known"), n=N_PREDS)
```
```{r}
# lets convert to a list and inspect it
pred1_list = as(pred1, "list")
length(pred1_list)
pred1_list[1]
```
Based on user 2 (user_id: kgRTizyE1oY441S8gn_eZw) pervious restaurant rating records, we could recommend this customer a list of 10 restauarants which someone having similar rating records like he/she also rated high star to.  

```{r}
pred1_list[2]
```

```{r}
calcPredictionAccuracy(pred1, 
                       getData(SAMPLE, "unknown"), 
                       goodRating = 5, 
                       given=N_PREDS)
```
True Positive : 0.0184
False Positive : 4.
