---
title: "Topic Modeling"
author: "Melissa Putur"
date: "4/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading the Libraries:
```{r message=FALSE, warning=FALSE}
library(sentimentr)
library(topicmodels)
library(tidyverse)
library(quanteda)
library(topicmodels)
library(tidytext)
library(factoextra)
library(wordcloud)
library(quanteda)
library(tidyverse)
library(dplyr)
```

Loading and Reorganizing the Data:
```{r}
reviews <- read.csv("../Data/canada_reviews_ms.csv")
reviews$text <- as.character(reviews$text)
restaurants <- read.csv("../Data/canada_restaurants_ms_edits.csv")
restaurantprice <- restaurants %>% select(business_id, attributes_RestaurantsPriceRange2)
reviewsWPrice <- left_join(reviews, restaurantprice, by = "business_id")
reviewsWPrice <- reviewsWPrice %>% rename(resPrice = attributes_RestaurantsPriceRange2)

glimpse(reviewsWPrice)
```

To explore if customers at inexpensive restaurants value different aspects of their dining experience than customers at expensive restaurants we performed topic modeling. We attempted grouping restaurants and reviews in several different ways including comparing 1-star and 2-star reviews to 4 and 5 star reviews and comparing “\$” and “\$\$” restaurants to “\$\$\$” and “\$\$\$\$” restaurants. After exploring several subsets we found we were able to find the most unique topics between groups when we limited the dataset to only include “\$” (inexpensive) and “\$\$\$\$”(expensive) restaurants and 1-star and 5-star reviews written about those restaurants. 


\$\$\$\$ Restaurants Reviews written by customers that had a positive experience at expensive restaurants typically fall under two topics. The first topic is around “food preparation”. Customers that write about this topic commonly use phrases like “perfectly cooked”, “great food” and even use specific phrases about how the food was prepared like “medium rare”. The second topic focuses on “special occasions”. These customers use phrases like “tasting menu”, “special occasion”, and “first time” to describe their experience. These customers do not appear to regularly dine at expensive restaurants but are looking to be wowed with a special experience when they do. 

Topic Modeling for Expensive restaurants, 5-star reviews.
```{r}
sw <- c(stopwords(),"ã","brã","lã","e")
swPhrases <- c("ice_cream","bubble_tea","milk_tea","green_tea","noodle_soup","milk_tea","spring_rolls","fried_chicken","red_bean","foie_gras","caesar_salad","come_back","coconut_cream","cream_pie","coconut_cream_pie","duck_fat","highly_recommend","highly_recommended","holt_renfrew","jerk_chicken","banh_mi","pork_belly","hot_sauce","pulled_pork","chicken_shawarma","kensington_market","coffee_shop","ordered_chicken","burrito_boyz","pad_thai","one_star","never_come","never_go","coming_back","going_back","next_time","really_good","definitely_recommend","go_back","ruth_chris","make_sure","coming_back","recommend_place","never_going","even_though","butter_chicken","kimchi_fries","black_sesame","definitely_back","feel_like","felt_like","baked_goods")

fourDollarPositive <- reviewsWPrice %>%
  filter(stars %in% c(5)) %>% 
  filter(resPrice == 4)

nrow(fourDollarPositive)

pos_four_dollar_tokens = tokens(fourDollarPositive$text, 
                            remove_punct = T,
                            remove_numbers = T)

pos_four_dollar_tokens <- tokens_remove(pos_four_dollar_tokens, pattern = sw)
#ngrams = 2)

pos_four_sdfm = dfm(pos_four_dollar_tokens,
           remove = swPhrases,
           ngrams = 2:4) %>%
  dfm_trim(max_docfreq = .6,
           docfreq_type = "prop",
           min_termfreq = 5,
           termfreq_type = "count")

#topfeatures(pos_four_sdfm,25)

pos_four_dollar_dtm <- convert(pos_four_sdfm, "topicmodels") ###

pos_four_tm2 = LDA(pos_four_dollar_dtm, k = 2, control = list(seed = 888)) ###   ###
terms(pos_four_tm2, 15)

pos_four_tm3 = LDA(pos_four_dollar_dtm, k = 3, control = list(seed = 888)) ###   ###
terms(pos_four_tm3, 15)

pos_four_tm4 = LDA(pos_four_dollar_dtm, k = 4, control = list(seed = 888)) ###   ###
terms(pos_four_tm4, 15)

no_tm2beta <- tidy(pos_four_tm2, matrix = "beta")

no_tm2beta %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  arrange(-beta) %>% 
  ungroup() %>% 
  mutate(term = reorder_within(term,beta,topic)) %>% 
  ggplot(aes(term,beta)) +
  geom_col(fill = "#F15C4F" ) +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Topic Model, Positive Reviews, $$$$ Restaurants ") +
  theme_minimal()

```

When exploring 1 and 2-star reviews for expensive restaurants, unique topics were difficult to find. Top phrases in negative reviews included neutral phrases like “customer service”, “tasting menu”, and “first time”, however within the top fifteen phrases things like “minutes later”, “quality food”, and “food poisoning”.

Topic Modeling for Expensive restaurants, 1-star reviews.
```{r}
swPhrases <- c("ice_cream","bubble_tea","milk_tea","green_tea","noodle_soup","milk_tea","spring_rolls","fried_chicken","red_bean","foie_gras","caesar_salad","come_back","coconut_cream","cream_pie","coconut_cream_pie","duck_fat","highly_recommend","highly_recommended","holt_renfrew","jerk_chicken","banh_mi","pork_belly","hot_sauce","pulled_pork","chicken_shawarma","kensington_market","coffee_shop","ordered_chicken","burrito_boyz","pad_thai","one_star","never_come","never_go","coming_back","going_back","next_time","really_good","definitely_recommend","go_back","ruth_chris","make_sure","coming_back","recommend_place","never_going","even_though","butter_chicken","kimchi_fries","black_sesame","definitely_back","feel_like","felt_like","baked_goods","much_better","prime_rib", "steak_house","four_seasons")

fourDollarNegative <- reviewsWPrice %>%
  filter(stars %in% c(1)) %>% 
  filter(resPrice == 4)

neg_four_dollar_tokens = tokens(fourDollarNegative$text, 
                            remove_punct = T,
                            remove_numbers = T)

neg_four_dollar_tokens <- tokens_remove(neg_four_dollar_tokens, pattern = sw)
#ngrams = 2)

## dfm and combine two smartword sources
neg_four_sdfm = dfm(neg_four_dollar_tokens,
           remove = swPhrases,
           ngrams = 2:4) %>%
  dfm_trim(max_docfreq = .7,
           docfreq_type = "prop",
           #min_termfreq = 5,
           #termfreq_type = "count")
  )

#topfeatures(neg_four_sdfm,15)

neg_four_dollar_dtm <- convert(neg_four_sdfm, "topicmodels") ###

neg_four_tm2 = LDA(neg_four_dollar_dtm, k = 2, control = list(seed = 888)) ###   ###
terms(neg_four_tm2, 15)

neg_four_tm3 = LDA(neg_four_dollar_dtm, k = 3, control = list(seed = 888)) ###   ###
terms(neg_four_tm3, 15)

neg_four_tm4 = LDA(neg_four_dollar_dtm, k = 4, control = list(seed = 888)) ###   ###
terms(neg_four_tm4, 15)

neg_four_beta <- tidy(neg_four_tm2, matrix = "beta")

 neg_four_beta%>% 
  group_by(topic) %>% 
  top_n(15, beta) %>% 
  arrange(-beta) %>% 
  ungroup() %>% 
  mutate(term = reorder_within(term,beta,topic)) %>% 
  ggplot(aes(term,beta)) +
  geom_col(fill = "#F15C4F") +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Negative Four Beta") +
  theme_minimal()
```

\$ Restaurants: We found two core topics for positive reviews written about inexpensive restaurants. While “first time” is again a popular phrase, more patrons for these restaurants indicated that they regularly visit the establishment. The first topic can be labeled as “Consistent Food”. These customers use phrases like “can’t go wrong”, “every time” and “great food”. They are regulars at these restaurants and enjoy that they consistently have a good meal, no matter what they choose on the menu. The second topic can be labeled as “Friendly Service”. While these customers do enjoy and mention the food, they use phrases like “super friendly”, “staff friendly” and “friendly service” when giving five-star reviews indicating that a welcoming and accommodating staff sets the place apart. 
For negative reviews, customers discussed two topics, “Food” and “Service”. Comments around food that frequently came up when discussing a negative 


Topic Modeling for Inexpensive restaurants, 5-star reviews.
```{r}
swPhrases <- c("ice_cream","bubble_tea","milk_tea","green_tea","noodle_soup","milk_tea","spring_rolls","fried_chicken","red_bean","foie_gras","caesar_salad","come_back","coconut_cream","cream_pie","coconut_cream_pie","duck_fat","highly_recommend","highly_recommended","holt_renfrew","jerk_chicken","banh_mi","pork_belly","hot_sauce","pulled_pork","chicken_shawarma","kensington_market","coffee_shop","ordered_chicken","burrito_boyz","pad_thai","one_star","never_come","never_go","coming_back","going_back","next_time","really_good","definitely_recommend","go_back","ruth_chris","make_sure","coming_back","recommend_place","never_going","even_though","butter_chicken","kimchi_fries","black_sesame","definitely_back","feel_like",
"felt_like","baked_goods","much_better","prime_rib","steak_house","four_seasons","love_place","great_place","one_best","place_great","pretty_good","love_love","can_get","can_say","hk_milk_tea")
 
 
oneDollarPositive <- reviewsWPrice %>%
  filter(stars == 5) %>% 
  filter(resPrice == 1)

#nrow(oneDollarPositive)

one_dollar_tokens = tokens(oneDollarPositive$text, 
                           remove_punct = T,
                           remove_numbers = T)

one_dollar_tokens <- tokens_remove(one_dollar_tokens, pattern = sw)
#ngrams = 2)

## dfm and combine two smartword sources
sdfm = dfm(one_dollar_tokens,
           remove = swPhrases,
           ngrams = 2:4) %>%
  dfm_trim(max_docfreq = .6,
           docfreq_type = "prop",
           min_termfreq = 4,
           termfreq_type = "count")

#topfeatures(sdfm,20)

one_dollar_dtm <- convert(sdfm, "topicmodels") ###

one_tm2 = LDA(one_dollar_dtm, k = 2, control = list(seed = 888)) ###   ###
terms(one_tm2, 15)

one_tm3 = LDA(one_dollar_dtm, k = 3, control = list(seed = 888)) ###   ###
terms(one_tm3, 15)

one_tm4 = LDA(one_dollar_dtm, k = 4, control = list(seed = 888)) ###   ###
terms(one_tm4, 15)

no_tm2beta <- tidy(one_tm2, matrix = "beta")

no_tm2beta %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  arrange(-beta) %>% 
  ungroup() %>% 
  mutate(term = reorder_within(term,beta,topic)) %>% 
  ggplot(aes(term,beta)) +
  geom_col(fill = "#F15C4F" ) +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Topic Model, Positive Reviews, $ Restaurants ") +
  theme_minimal()

```

For negative reviews, customers discussed two topics, “Food” and “Service”. Comments around food that frequently came up when discussing a negative experience with food were comments regarding bad taste as well as the fact that the food may have given them food poisoning. Customers who discussed service mentioned things like “minutes later” and “behind the counter”. These customers were dissatisfied with the length of time it took their food to be ready. Also, a problem that is more unique to inexpensive vs. expensive restaurants is that customers are able to watch employees behind the counter interact and/or prepare their food. Customers with a negative experience often discussed behavior they saw behind the counter that detracted from their experience. Finally, these customers more frequently used phrases like “came back” and “last time” indicating that it was not their first time dining at the restaurant. 


Topic Modeling for Inexpensive restaurants, 1-star reviews. 
```{r}
swPhrases <- c("ice_cream","bubble_tea","milk_tea","green_tea","noodle_soup","milk_tea","spring_rolls","fried_chicken","red_bean","foie_gras","caesar_salad","come_back","coconut_cream","cream_pie","coconut_cream_pie","duck_fat","highly_recommend","highly_recommended","holt_renfrew","jerk_chicken","banh_mi","pork_belly","hot_sauce","pulled_pork","chicken_shawarma","kensington_market","coffee_shop","ordered_chicken","burrito_boyz","pad_thai","one_star","never_come","never_go","coming_back","going_back","next_time","really_good","definitely_recommend","go_back","ruth_chris","make_sure","coming_back","recommend_place","never_going","even_though","butter_chicken","kimchi_fries","black_sesame","definitely_back","feel_like", "felt_like","baked_goods","much_better","prime_rib", "steak_house","four_seasons", "love_place","great_place","one_best","place_great","pretty_good","love_love","can_get","can_say","hk_milk_tea", "avoid_place","never_coming","stay_away","give_place", "never_go_back","last_night")

oneDollarNegative <- reviewsWPrice %>%
  filter(stars == 1) %>% 
  filter(resPrice == 1)

#nrow(oneDollarNegative)

one_dollar_tokens = tokens(oneDollarNegative$text, 
                           remove_punct = T,
                           remove_numbers = T)

one_dollar_tokens <- tokens_remove(one_dollar_tokens, pattern = sw)
#ngrams = 2)

## dfm and combine two smartword sources
sdfm = dfm(one_dollar_tokens,
           remove = swPhrases,
           ngrams = 2:4) %>%
  dfm_trim(max_docfreq = .6,
           docfreq_type = "prop",
           min_termfreq = 5,
           termfreq_type = "count")

topfeatures(sdfm,15)

one_dollar_dtm <- convert(sdfm, "topicmodels") ###

one_tm2 = LDA(one_dollar_dtm, k = 2, control = list(seed = 888)) ###   ###

no_tm2beta <- tidy(one_tm2, matrix = "beta")

no_tm2beta %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  arrange(-beta) %>% 
  ungroup() %>% 
  mutate(term = reorder_within(term,beta,topic)) %>% 
  ggplot(aes(term,beta)) +
  geom_col(fill = "#F15C4F" ) +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Topic Model, Negative Reviews, $ Restaurants ") +
  theme_minimal()
```

