---
title: "Project Overview - Can Yelp Reviews Save the Restaurant Industry?"
author: "Xinman Liu, Melissa Putur, Jiao Sun, Adil Bin Wahab, Rui Xu"
date: "2/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(jsonlite)
library(tidytext)
library(stringr)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(readr)
library(tidyr)
library(lubridate)
library(gridExtra)
library(quanteda)
```

```{r message=FALSE, warning=FALSE}
#Loading the datasets
#restaurants <- read_csv("Data/canada_restaurants_ms.csv") #missy file path
restaurants <- read_csv("Data/canada_restaurants_ms_edits.csv")
#reviews <- read_csv("/Users/Adil/Documents/RStudio/capstoneData/canada_reviews.csv") #adil file path
reviews <- read_csv("Data/canada_reviews_ms.csv")
#checkIns <- read_csv("/Users/Adil/Documents/RStudio/capstoneData/canada_checkin.csv")
checkin <- read_csv("Data/canada_checkins_ms.csv")
viewers <- read_csv("Data/canada_viewers_rui_edits.csv")
```

### Business Problem   
Finding restaurants through directory services apps and review forum apps have become a crucial part of business discovery. It has changed the way restaurants of different tiers attain brand loyalty and market themselves.   

Yelp’s terabytes of historical data on restaurants’ reviews, check-ins, and special attributes have led to the ability to uncover valuable trends to aid businesses in understanding their impact on customers and what make customers tick.   

How can a restaurant leverage information on Yelp to increase their chances of attaining their business goals and preventing permanent closure? What aspects make customers choose a restaurant over its competitors? 


### Ways to Solve Our Problem
**1. descriptive + diagnostic analytics**
Study the most essential aspects of a restaurant based on cuisine and price level to understand what most customers value (e.g. Do customers at low-price restaurants value the same “things” as mid-price and high-price restaurants).
**2. predictive analytics**
Predict whether a restaurant will close in the near future based on specific variables.
**3. prescriptive Analytics**
Create a recommendation system that predicts users’ preferences based on their reviews. prescriptive Analytics


### Data Source   
We got our data from [Kaggle](https://www.kaggle.com/yelp-dataset/yelp-dataset), and there are 4 datasets in total, namely, business, check-in, reviews, and users. 
**Business dataset** includes business types, attributes, location, review count, star ratings, and open hours.
**User dataset** about reviews on each business: text, rating, review count, and reactions (useful, cool, funny).
**Check-ins datasets** includes the exact timestamp a user checked into a business.
**Review dataset** includes the review text of each business with user ids, usefulness count, and reactions (useful, cool, funny).


### Data Cleaning
**Business Dataset:** For the project we are only interested in exploring reviews, star ratings, and interactions with restaurants. The business dataset contains information for all business reviewed by yelp. To narrow down this dataset we manually reviewed the top 1,000 phrases that businesses were tagged with to identify phrases that would be a good fit for restaurant. We filtered the complete list to only include businesses that were tagged with the phrases “food” and/or “restaurant”. Using these phrases we were able to identify ~75,000 restaurants. 
**Review Dataset:** The review dataset contains reviews for all businesses on Yelp. We reduced the review dataset to only include reviews for restaurants (as identified by the above processing steps). After this reduction the dataset contained over 4 million reviews.
**Check-In Dataset:** The check-in dataset was structured so that each business was represented as a row with a comma separated list of check-ins by date. To make the data easier to work with we rearranged it so that each check-in was represented as a row rather than a list.
**Final Data Cleaning Criteria:** To balance having a robust datset that minimizes long compute time while preserving data quality, the dataset was filtered to only include Canadian cities (Toronto, Mississauga and Calgary). Montréal was excluded as a majority of the reviews are in French. Furthermore to ensure rich, non-sparse data for for the remainder of restaurants in the dataset, we futher reduced the restaurant list to include only the restaurants with the top 50% of reviews. With this criteria only restaurants with 14+ reviews remained. 
**Post-cleaned Dataset Information**: 
Table 1: Number of reviews of each resturant (`nrow(reviews)` rows) - 200 KB (CSV)
Table 2: Restaurant general info and attributes (`nrow(restaurants)` rows) - reduced to 4 MB (CSV) from 34 MB (JSON)
Table 3: Restuarant check-ins (`nrow(checkin)` rows) - reduced to 38.5 MB (CSV) from 390 MB (JSON)
Table 4: Restaurant reviews (371,610 rows) - reduced to 320 MB (CSV) from 4.98 GB (JSON)

###Exploratory Data Analysis

The dataset contained 7,530 restaurants from three Canadian cities. Toronto had the majority of restaurants 5,372 (70%), followed by Calgary with 1,365 (18%), and Mississauga with 794 (11%). The dataset contained both open and permanently closed restaurants; 5,699 (76%) restaurants were open and 1,831 (24%) restaurants had gone out of business.
```{r}
fig1 <- restaurants %>% 
  ggplot(aes(x=city))+
  geom_bar(fill=c("#F15C4F", "#FCD6D3","#F8ADA8"))+
  theme_classic()+
  labs(title = "Restaurant Distribution by City (Fig. 1)",
       y="# of Restaurants") +
  theme(axis.text = element_text(size = 13)) +
  theme(axis.title = element_text(size = 13)) 

open <- restaurants %>% group_by(is_open) %>% count(is_open)
open[open$is_open==0,1]="no"
open[open$is_open==1,1]="yes"
fig2 <- open %>% 
  ggplot(aes(is_open,n))+
  geom_col(fill=c("#F15C4F", "#F8ADA8"))+
  labs(title = "Current Restuarant Status (Fig. 2)",
       x="Still in Business?",
       y="# of Restaurants")+
  theme_classic() +
   theme(axis.text = element_text(size = 13)) +
  theme(axis.title = element_text(size = 13)) 

fig1
```

```{r}
fig2
```


Check-In dates ranged from 2010 to 2018. The average restaurant had 111 check-ins with a median of 57 and a maximum of 2,872 check-ins. Check-ins steadily increased year over year in the data until about 2015-2017 when they reached a peak. The first year to see a decrease in check-ins from the previous year was 2018. The check-in data does exhibit some seasonality with visits appearing to typically be higher in the summer months than winter months. 

```{r message=FALSE, warning=FALSE}
fig3 <- checkin %>% 
  group_by(updatedcheckins) %>% 
  add_count() %>% 
  ggplot() +
  geom_smooth(aes(x = month(updatedcheckins), y = n, color = as.factor(year(updatedcheckins))), size = 2, se = F) + 
  theme_minimal()+
  theme(axis.text.x=element_text(angle=90, hjust=1)) +
  labs(title = "Check-In by Day (Fig. 3)", x = "month", y="count", color = "year") +
  scale_x_continuous(breaks=seq(0,12,1)) +
  scale_colour_manual(values = c("#D32323","#F15C4F","#F8ADA8","#FCD6D3","#E6E6E6","#999999","#CCCCCC","#666666","#4D4D4D","#333333"))

fig4 <- checkin %>% 
  group_by(business_id) %>% 
  count(.) %>% 
  ggplot() +
  geom_histogram(aes(x = n), fill="#D32323", binwidth = 10) +
  labs(title="Distribution of Check-In Count (Fig. 4)", x = "Check-Ins")+
  theme_classic()

fig3
```
```{r}
fig4
```


The distribution of star ratings for restaurants was left skewed. The majority of restaurants had a four-star rating with the second most common rating being 3.5 stars. The distribution of star ratings between the three cities was fairly similar, however, Calgary had a slightly higher percentage of 4+ star restaurants than Toronto and Mississauga. 

```{r message=FALSE, warning=FALSE}
fig5 <- restaurants %>% 
  ggplot() +
  geom_bar(aes(stars),fill="#F15C4F") +
  labs(title = "Distribution of Star Ratings (Fig. 5)", x = "Star Rating", y = "# of Restaurants") +
  theme_minimal()
fig5
```


```{r}
fig6 <- restaurants %>% 
  count(city, stars) %>% 
  group_by(city) %>% 
  mutate(percent = n / sum(n) ) %>% 
  ungroup() %>% 
  ggplot(aes(x = city, y = percent, fill = as.factor(stars))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title="Distribution of Star Ratings by City (Fig. 6)", x = "City", y = "Percent of Restaurants", fill = "Star Rating") +
  scale_fill_manual(values = c("#D32323", "#F15C4F", "#F8ADA8", "#FCD6D3", "#E6E6E6", "#999999", "#CCCCCC","#666666","#4D4D4D","#333333")) +
  theme_minimal() +
  theme(axis.text = element_text(size = 13)) +
  theme(axis.title = element_text(size = 13)) 
fig6
```

The dataset contained over 480k reviews. The mean number of reviews per restaurant was 64. The distribution of the review length is right-skewed and more than 50% of all the reviews were between 25 to 120 words. The average review length was around 130 words while the median was around 100 words. The most common bigrams found in reviews included “come back”, “ice cream” and “go back”. 

```{r}
print(paste0("number of reviews: ", nrow(reviews)))
print(paste0("average # of reviews: ", round(nrow(reviews)/nrow(restaurants))))
```

```{r}
stringsAsFactors = FALSE
review_length <- sapply(strsplit(as.character(reviews$text), " "), length)
review_length <- as.data.frame(review_length)

# Dist freq 
fig7 <- ggplot(data=review_length, aes(x=review_length)) +
geom_histogram(aes(y = stat(count) / sum(count)),
               bins = 25, fill="#D32323") +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Distribution of Review Length (Fig. 7)",
       x="Review Length",y="Relative Frequency") +
  theme_minimal() + scale_x_continuous(breaks=seq(0,1000,25)) +
  theme(axis.text.x=element_text(angle=90, hjust=1))

fig7
```

```{r}
tokens = tokens(as.character(reviews$text), 
                remove_punct = T,
                remove_numbers = T)

sw <- c(stopwords(),"ã","brã","lã","e")
tokens <- tokens_remove(tokens, pattern = sw)
sdfm = dfm(tokens,
           ngrams = 2)
topBigrams <- as.data.frame(topfeatures(sdfm,10))
names(topBigrams)[1] <- "Count"
bigrams <- rownames(topBigrams)
topBigrams$Bigrams <- bigrams

fig8 <- topBigrams %>% 
  arrange(desc(Count)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(Bigrams, Count), y = Count), fill = "#F15C4F") +
  coord_flip() +
  theme_minimal() +
  labs(title="Customer Review Top Bigrams (Fig. 8)", x = "Bigram", y = "Count") +
  theme(axis.text = element_text(size = 13)) +
  theme(axis.title = element_text(size = 13))
fig8
```

Reviewers submitting 1 and 2-star reviews wrote longer reviews on average than reviewers who gave 3, 4, and 5-star reviews. The number of reviews increased year over year to a peak in 2017 of just under 100,000 reviews.

```{r}
reviews$review_length <- review_length$review_length

fig9 <- reviews %>%
  group_by(stars) %>% 
  summarise(avgLength = round(mean(review_length))) %>% 
  arrange(desc(avgLength)) %>% 
  ggplot() +
  geom_col(aes(x = stars, y = avgLength), fill = "#F8ADA8") +
  labs(title="Star Rating by Review Length (Fig.9)",
       x="Star Rating",y="Avg Review Length (words)") +
  theme_minimal() +
  theme(axis.text = element_text(size = 13)) +
  theme(axis.title = element_text(size = 13)) 
fig9
```

```{r}
posts = reviews %>%
  select(date, review_count, user_id) %>%
  mutate(date = year(date)) %>%
  rename(year = date) %>%
  count(year) %>%
  mutate(cum = cumsum(n))

fig10 <- posts %>%
  ggplot() +
  geom_bar(aes(x = year, y = n), stat = 'identity', fill = "#f8ada8") +
  geom_line(aes(x = year, y = cum)) +
  scale_x_continuous(breaks = reviewer$year) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
  labs(title = "Number of Reviews by Year (Fig. 10)",
       subtitle = "and accumulated reviews count",
       x = 'Years',
       y = 'Number of reviews')+
  theme_minimal()
fig10
```

By 2018, there were over 900,000 users who had written at least one review, or checked-in once at one of the restaurants in the dataset. The number of new users increased year over year until 2015 when annual new users peaked. 

For the purposes of this project, an active user is defined as someone who posts more than the average reviewer in a year. The number of users who are considered active was increasing year over year until 2018. The percent of users who were considered active in 2017 was 24% compared to 22% in 2018. Although the total number of users grew year over year, the number of reviews written on average steadily decreased from a peak of around 7 reviews per user in 2007 to just over 2 reviews in 2018. 

```{r}
data = reviews %>%
  select(yelping_since, review_count, user_id)
data = data[!duplicated(data),]
reviewer = data %>%
  mutate(year = year(yelping_since)) %>%
  count(year) %>%
  mutate(cum = cumsum(n))

fig11 <- reviewer %>%
  ggplot() +
  geom_bar(aes(x = year, y = cum), stat = 'identity', fill = "#F15C4f") +
  geom_line(aes(x = year, y = n)) +
  scale_x_continuous(breaks = reviewer$year) +
  labs(title = "Frowth of Active Users on Yelp (Fig. 11)",
       subtitle = "and annual increase amount",
       x = 'Years',
       y = 'Number of Reviewers') +
  theme_minimal()
fig11
```

For the purposes of this project, an active user is defined as someone who posts more than the average reviewer in a year. The number of users who are considered active was increasing year over year until 2018. The percent of users who were considered active in 2017 was 24% compared to 22% in 2018. Although the total number of users grew year over year, the number of reviews written on average steadily decreased from a peak of around 7 reviews per user in 2007 to just over 2 reviews in 2018. 
```{r message=FALSE, warning=FALSE}
mean_n = reviews %>% 
  mutate(year = year(date)) %>%
  group_by(year, user_id) %>%
  count(year) %>%
  group_by(year) %>%
  summarise(mean = mean(n)) 

# Label active y/n
tmp = reviews %>%
  mutate(year = year(date)) %>%
  group_by(year, user_id) %>%
  count(year)
active = merge(tmp, mean_n, by = 'year')
act_year = active %>% 
  group_by(year) %>%
  mutate(active = ifelse(n > mean, 1, 0)) %>%
  mutate(active = factor(active)) 

gf = act_year %>%
  ungroup() %>%
  group_by(year, mean, active) %>%
  count(active)

fig12 <- ggplot(gf,aes(x=year)) + 
  geom_bar(aes(y=n, fill = active), stat = 'identity', alpha = 0.8) +
  scale_fill_manual(values= c("#f8ada8", "#F15C4f")) +
  scale_x_continuous(breaks = gf$year) + 
  labs(y='Users',
       title = 'Yelp active users gradually abate from 2007 to 2018 (Fig. 12)') +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent",colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA)
  ) + 
  geom_line(aes(y = mean*4000)) + 
  scale_y_continuous(sec.axis = sec_axis(~./4000, name = "Average num of review posted"))
fig12

```

Not only did active users write more reviews on average, but the length of their reviews were also longer. The average active reviewer wrote reviews containing just under 125 words compared to the average inactive user whose reviews typically contained under 95 words. 

Active and inactive users gave different star ratings on average. Active users rate in a more normal way with a peak at about four stars, while inactive users are more extreme in their rating. Inactive users give 1-star and 5-star reviews more often than active users. 

```{r}
# 2018 reviewers
active18 = act_year %>%
  filter(year == 2018) %>%
  ungroup() %>%
  mutate(year_act = 2018) %>%
  select(user_id, annual_reviews = n, active, year_act) 

viewer18 = inner_join(viewers, active18, by='user_id')
reviews18 = inner_join(viewer18, reviews, by = 'user_id')

# 2017 reviewers
active17 = act_year %>%
  filter(year == 2017) %>%
  ungroup() %>%
  mutate(year_act = 2017) %>%
  select(user_id, annual_reviews = n, active, year_act)
viewer17 = inner_join(viewers, active17, by='user_id')
reviews17 = inner_join(viewer17, reviews, by = 'user_id')

# 2016 reviewers
active16 = act_year %>%
  filter(year == 2016) %>%
  ungroup() %>%
  mutate(year_act = 2016) %>%
  select(user_id, annual_reviews = n, active, year_act)
viewer16 = inner_join(viewers, active16, by='user_id')
reviews16 = inner_join(viewer16, reviews, by = 'user_id')

# 2015 reviewers
active15 = act_year %>%
  filter(year == 2015) %>%
  ungroup() %>%
  mutate(year_act = 2015) %>%
  select(user_id, annual_reviews = n, active, year_act)
viewer15 = inner_join(viewers, active15, by='user_id')
reviews15 = inner_join(viewer15, reviews, by = 'user_id')

## combine
reviewers15_18 = rbind(reviews15, reviews16, reviews17, reviews18)

viewers15_18 = reviewers15_18 %>%
  mutate(year_post = year(date)) %>%
  filter(year_post == year_act) %>%
  select(user_id, avg_stars = average_stars.x, friends, useful, fans, 
         yelp_since = yelping_since.x, 
         total_reviews = review_count.x,
         annual_reviews, active, year_act, text, date, stars, business_id)
## active ~ star rating
tmp_stars = viewers15_18 %>%
  select(user_id, stars, active, year_act) %>%
  group_by(user_id, active, year_act) %>%
  summarise(avg_annual_stars = round(mean(stars), 2)) %>%
  ungroup() %>%
  mutate(active = ifelse(active==1, 'yes', 'no')) 

fig13 <- ggplot(tmp_stars, aes(avg_annual_stars, fill = active)) + 
  scale_fill_manual(values= c("#626262", "#F15C4f")) +
  geom_density(alpha = 0.2) +
  facet_wrap(.~ year_act, ncol = 2) +
  labs(x='Annual rating in average',
       title = 'Active vs. Inactive Rating Patterns (Fig. 13)') +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent",colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA)
  )
fig13
```

```{r}
## active ~ word.length
viewers15_18 = viewers15_18 %>%
  mutate(words_length = sapply(strsplit(text, " "), length))

words_length = viewers15_18 %>%
  select(user_id, words_length, active, year_act) %>%
  group_by(year_act, active) %>%
  summarise(avg.words = mean(words_length))

fig14 <- ggplot(words_length, aes(x=year_act, y=avg.words)) +
  geom_line(aes(color=active)) +  
  scale_color_manual(values= c("#f8ada8", "#F15C4f")) +
  labs(x='Active reviewers',
       y='Average word length in reviews',
       title = 'Active reviewers write longer reviews (Fig. 14)') +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent",colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA)
  ) 
fig14
```

