---
title: "Random Forest"
author: "Jiao Sun"
date: "5/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(zoo)
library(DescTools)
library(lubridate)
library(randomForest)
library(gbm)
```


```{r}
restaurants_lm <- read_csv("canada_restaurants_ms.csv")
checkIns <- read_csv("canada_checkins_ms.csv")
```

Data cleaning
```{r}
# Check NA count of columns
View(as.data.frame(colSums(is.na(restaurants_lm))))
# attributes_BusinessAcceptsCreditCards	7158  -- too many NAs to keep col -- DELETE
# attributes_Caters	2296  -- over 30% NAs, not worth keeping -- DELETE
# attributes_NoiseLevel	1592
# attributes_Alcohol	1436
# attributes_HasTV	1335
# attributes_RestaurantsAttire	1329  -- make 0.None 1.casual 2.dressy 3.formal

# attributes_Ambience	1296  -- {x:y} format
unique(unlist(str_split(restaurants_lm$attributes_Ambience, ' ' )))
## casual - trendy - hipster - divey - touristy - romantic - intimate - classy - upscale

# attributes_WiFi	1267  -- 0.no/none 1.paid 2.free
# attributes_RestaurantsReservations	1196
# attributes_GoodForKids	1193
# attributes_RestaurantsGoodForGroups	1126
# attributes_RestaurantsDelivery	1074
# attributes_BikeParking	1005

# attributes_BusinessParking	390 --{x:y} format
unique(unlist(str_split(restaurants_lm$attributes_BusinessParking, ' ' )))
## garage - street - validated - lot - valet 

# Turn unique ambience values into columns
restaurants_lm$amb_casual <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'casual': True"), 1, 0)
restaurants_lm$amb_trendy <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'trendy': True"), 1, 0)
restaurants_lm$amb_hipster <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'hipster': True"), 1, 0)
restaurants_lm$amb_touristy <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'touristy': True"), 1, 0)
restaurants_lm$amb_romantic <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'romantic': True"), 1, 0)
restaurants_lm$amb_intimate <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'intimate': True"), 1, 0)
restaurants_lm$amb_classy <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'classy': True"), 1, 0)
restaurants_lm$amb_upscale <- ifelse(str_detect(restaurants_lm$attributes_Ambience, "'upscale': True"), 1, 0)

# Turn 5 different parking options into separate columns and create dummies
restaurants_lm$park_garage <- ifelse(str_detect(restaurants_lm$attributes_BusinessParking, "'garage': True"), 1, 0)
restaurants_lm$park_street <- ifelse(str_detect(restaurants_lm$attributes_BusinessParking, "'street': True"), 1, 0)
restaurants_lm$park_validated <- ifelse(str_detect(restaurants_lm$attributes_BusinessParking, "'validated': True"), 1, 0)
restaurants_lm$park_lot <- ifelse(str_detect(restaurants_lm$attributes_BusinessParking, "'lot': True"), 1, 0)
restaurants_lm$park_valet <- ifelse(str_detect(restaurants_lm$attributes_BusinessParking, "'valet': True"), 1, 0)

# Make dummies for NoiseLevel
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="None"] <- 0
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="'quiet'"] <- 1
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="u'quiet'"] <- 1
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="u'average'"] <- 2
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="'average'"] <- 2
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="'loud'"] <- 3
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="u'loud'"] <- 3
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="u'very_loud'"] <- 4
restaurants_lm$attributes_NoiseLevel[restaurants_lm$attributes_NoiseLevel=="'very_loud'"] <- 4
unique(restaurants_lm$attributes_NoiseLevel)

# Make dummies for Alcohol 
restaurants_lm$attributes_Alcohol[restaurants_lm$attributes_Alcohol=="u'full_bar'"]="full_bar"
restaurants_lm$attributes_Alcohol[restaurants_lm$attributes_Alcohol=="'full_bar'"]="full_bar"
restaurants_lm$attributes_Alcohol[restaurants_lm$attributes_Alcohol=="u'beer_and_wine'"]="beer_and_wine"
restaurants_lm$attributes_Alcohol[restaurants_lm$attributes_Alcohol=="'beer_and_wine'"]="beer_and_wine"
restaurants_lm$attributes_Alcohol[restaurants_lm$attributes_Alcohol=="'none'"]="None"
restaurants_lm$attributes_Alcohol[restaurants_lm$attributes_Alcohol=="u'none'"]="None"
unique(restaurants_lm$attributes_Alcohol)
# Make two columns from alcohol attribute: full_bar and beer_and_wine
restaurants_lm$full_bar <- ifelse(restaurants_lm$attributes_Alcohol=="full_bar", 1, 0)
restaurants_lm$beer_wine <- ifelse(restaurants_lm$attributes_Alcohol=="beer_and_wine", 1, 0)

# Make dummies for RestaurantsAttire: none = 0; casual = 1; dressy = 2; formal = 3
restaurants_lm$attributes_RestaurantsAttire[restaurants_lm$attributes_RestaurantsAttire=="None"] <- 0
restaurants_lm$attributes_RestaurantsAttire[restaurants_lm$attributes_RestaurantsAttire=="'casual'"] <- 1
restaurants_lm$attributes_RestaurantsAttire[restaurants_lm$attributes_RestaurantsAttire=="u'casual'"] <- 1
restaurants_lm$attributes_RestaurantsAttire[restaurants_lm$attributes_RestaurantsAttire=="'dressy'"] <- 2
restaurants_lm$attributes_RestaurantsAttire[restaurants_lm$attributes_RestaurantsAttire=="u'dressy'"] <- 2
restaurants_lm$attributes_RestaurantsAttire[restaurants_lm$attributes_RestaurantsAttire=="'formal'"] <- 3
restaurants_lm$attributes_RestaurantsAttire[restaurants_lm$attributes_RestaurantsAttire=="u'formal'"] <- 3

# Make dummies for WiFi: none/no = 0; paid = 1; free = 2
restaurants_lm$attributes_WiFi[restaurants_lm$attributes_WiFi=="None"] <- 0
restaurants_lm$attributes_WiFi[restaurants_lm$attributes_WiFi=="u'no'"] <- 0
restaurants_lm$attributes_WiFi[restaurants_lm$attributes_WiFi=="'no'"] <- 0
restaurants_lm$attributes_WiFi[restaurants_lm$attributes_WiFi=="u'paid'"] <- 1
restaurants_lm$attributes_WiFi[restaurants_lm$attributes_WiFi=="'paid'"] <- 1
restaurants_lm$attributes_WiFi[restaurants_lm$attributes_WiFi=="u'free'"] <- 2
restaurants_lm$attributes_WiFi[restaurants_lm$attributes_WiFi=="'free'"] <- 2

# Create month columns for checkins 
checkIns %>% group_by(business_id) %>% 
  mutate(j=ifelse(month(updatedcheckins)==1,1,0)) %>%
  mutate(f=ifelse(month(updatedcheckins)==2,1,0)) %>%
  mutate(mr=ifelse(month(updatedcheckins)==3,1,0)) %>%
  mutate(ap=ifelse(month(updatedcheckins)==4,1,0)) %>%
  mutate(m=ifelse(month(updatedcheckins)==5,1,0)) %>%
  mutate(ju=ifelse(month(updatedcheckins)==6,1,0)) %>%
  mutate(jl=ifelse(month(updatedcheckins)==7,1,0)) %>%
  mutate(au=ifelse(month(updatedcheckins)==8,1,0)) %>%
  mutate(s=ifelse(month(updatedcheckins)==9,1,0)) %>%
  mutate(o=ifelse(month(updatedcheckins)==10,1,0)) %>%
  mutate(n=ifelse(month(updatedcheckins)==11,1,0)) %>%
  mutate(d=ifelse(month(updatedcheckins)==12,1,0)) %>%
  summarize(jan=sum(j),feb=sum(f),mar=sum(mr),apr=sum(ap),may=sum(m),jun=sum(ju),
            jul=sum(jl),aug=sum(au),sep=sum(s),oct=sum(o),nov=sum(n),dec=sum(d)) -> checkIn_months

# check-ins by season w/ avg per month 
checkIn_months %>% group_by(business_id) %>% 
  mutate(winter = sum(mean(dec),mean(jan),mean(feb)), spring = sum(mean(mar),mean(apr),mean(may)), 
         summer = sum(mean(jun),mean(jul),mean(aug)), fall = sum(mean(sep),mean(oct),mean(nov))) %>% 
  select(winter, spring, summer, fall) -> checkIn_seasons

# Make new df w/ columns for ML application -- ml_df
restaurants_lm %>% 
  mutate(attributes_GoodForKids=1*attributes_GoodForKids,
         attributes_RestaurantsReservations=1*attributes_RestaurantsReservations,
         attributes_RestaurantsTakeOut=1*attributes_RestaurantsTakeOut,
         attributes_OutdoorSeating=1*attributes_OutdoorSeating,
         attributes_HasTV=1*attributes_HasTV,
         attributes_RestaurantsGoodForGroups=1*attributes_RestaurantsGoodForGroups,
         attributes_RestaurantsDelivery=1*attributes_RestaurantsDelivery,
         attributes_BikeParking=1*attributes_BikeParking,
         attributes_NoiseLevel=as.numeric(attributes_NoiseLevel),
         attributes_WiFi=as.numeric(attributes_WiFi),
         attributes_RestaurantsAttire=as.numeric(attributes_RestaurantsAttire),
         price_range=attributes_RestaurantsPriceRange2,
         review_count=Winsorize(review_count, probs = c(0.05, 0.95))) %>%
  # Replace NAs with column median
  mutate_all(~ifelse(is.na(.), median(., na.rm = TRUE), .)) %>% 
  select(business_id, is_open, price_range, stars, review_count,
         attributes_RestaurantsReservations, attributes_RestaurantsTakeOut,
         attributes_RestaurantsDelivery, attributes_HasTV, attributes_WiFi,
         amb_casual, amb_trendy, amb_hipster, amb_touristy,
         amb_romantic, amb_intimate, amb_classy, amb_upscale,
         full_bar, beer_wine, attributes_RestaurantsAttire,
         attributes_GoodForKids, attributes_RestaurantsGoodForGroups,
         attributes_BikeParking, attributes_OutdoorSeating,
         park_street, park_validated, park_lot, park_garage, park_valet
  ) -> ml_df

ml_df %>% left_join(checkIn_seasons,by='business_id') %>% 
  select(-business_id) -> ml_df
ml_df <- na.omit(ml_df)
```

Spliting Data For Train/Test
```{r}
set.seed(007)

# Create formula where y = is+_open
formula_isOpen <- as.formula(is_open ~ ., ml_df)

# Split train (70%) and test (30%)
train <- round(0.7 * nrow(ml_df))
test <- nrow(ml_df) - train

train_index <- sample(nrow(ml_df), train) # index of random rows 

# Train and test data 
data_train <- ml_df[train_index,]
data_test <- ml_df[-train_index,]

# Split Xs and Ys for Lasso, Random Forest, and Ridge
x_train_isOpen <- model.matrix(formula_isOpen, data_train)[, -1]
x_test_isOpen <- model.matrix(formula_isOpen, data_test)[, -1]

y_train_isOpen <- data_train$is_open
y_test_isOpen <- data_test$is_open
```

Random Forest
```{r}
xnames_rf=colnames(ml_df)
xnames_rf = xnames_rf[!xnames_rf %in% c("is_open", "review_count")]

lf <- "is_open ~ price_range"
for (i in 2:length(xnames_rf)) {
  as.character(xnames_rf[i])
  lf <- paste(lf, "+", xnames_rf[i], sep = " ")
}
f <- as.formula(lf)
x_train <- model.matrix(f,data_train)[ ,-1]
y_train <- data_train$is_open
x_test <- model.matrix(f,data_test)[ , -1]
y_test <- data_test[ ,"is_open"]

x_train_isOpen <- model.matrix(formula_isOpen, data_train)[, -1]
x_test_isOpen <- model.matrix(formula_isOpen, data_test)[, -1]

y_train_isOpen <- data_train$is_open
y_test_isOpen <- data_test$is_open

fit_rf <- randomForest(f,
                       data_train,
                       ntree=500,
                       do.trace=F)

varImpPlot(fit_rf)

predicted <- predict(fit_rf, data_test, type="response")
library(InformationValue)
plotROC(data_test$is_open, predicted)
confusionMatrix(data_test$is_open, predicted)
```

The final random forest model included 30 variables. Using the variable importance plot we can see that the variables related to the attributes of the restaurant are the most predictive. On the plot we can see "summer", "fall", "spring", "winter", "stars", "price_range", are the most predictive with IncNodepurity all higher than 60. We got the accurcy rate for this model is almost 70.3%.